Solving TSP With Genetic Algorithms

### 2018-04-13 tex/paper.tex
Added references and table for results

\usepackage{bm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}

\begin{document}	
	\twocolumn[
	\begin{@twocolumnfalse}
		\author{Mohamed Moanis Ali, Michael Sedrak \\  mohamed.moanis.ali@gmail.com, m.sedrak@live.com}
		\title{Solving TSP With Genetic Algorithms}
		\maketitle
		\begin{abstract}
			The traveling salesman problem is considered one of hardest problems in computer science with no polynomial runtime algorithm to solve it so far. Nevertheless, TSP is an active research point with more algorithms being developed to either approximately solve the original problem, or to speed it up so larger datasets can be solved. In this work we present a comparative study using four algorithms to solve TSP -genetic algorithm, simulated annealing, nearest neighbor and ant colony optimization to analyze the runtime performance and the correctness of their solution.  The results of the experiments on the Uruguay dataset -using both {Matlab\texttrademark}  and {Octave\texttrademark}- showed that the nearest neighbor algorithm provided the fastest runtime and a relatively correct solution with $18\%$ from the known optimal solution, while the ant colony optimization proved to be the least correct among the studied algorithms. The study also found a huge performance difference between {Matlab\texttrademark}  and {Octave\texttrademark} software.
		\end{abstract}

	\end{@twocolumnfalse}
	]

@@

	\subsection{Nearest Neighbor (NN)}
	Nearest neighbor approach is one of the simplest algorithms in combinatorial optimization problems and also one of the fastest running algorithms.

@@

	For better understanding of the ACO algorithms, an example from \textbf{[\underline{link}]} is illustrated. As shown in Fig.  \ref{fig:ACOexample} a, the system is shown with the initial start points A as the food source ,and E as the colony nest, and the distances of the paths. At t=0, 30 ants start moving from both E and A simultaneously. The ants currently can see no preference between the different paths because there are no traces of pheromone left behind, therefore, the ants divide randomly with equal probability going in to both points H and C, i.e. 15 ants are going from D to C and 15 ants from D to H, and similarly from point B to both C and H. At t=1, the 30 new ants coming from B, find 2 unequal paths, the path going to H has a trail of 15 from the ants that went to H from B, while the path going to C has a trail of 30 that can be divided into:
	\begin{enumerate}
		\item The trail from ants going from point B.
		\item The trail from ants coming from point D.
	\end{enumerate}
	 Resulting in a bias in the probability of choosing the path unlike the previous step. The number of ants going to C is expected to be double the number of ants going to H. The same is true for the 30 new ants coming from D.
	\begin{figure}[h!]
		\centering
		\includegraphics[scale=0.6]{./ACOexample.jpg}
		\caption{ACO example \textbf{[\underline{link}]} (a)The initial description of the problem showing distances. (b)At time t=0 no trail is found on the graph therefore ants choose the paths with equal probability. (c)Trails from previous step are found and therefore the choice of paths is biased towards the shorter edges.}
		\label{fig:ACOexample}
	\end{figure}

	In combinatorial problems, we deal with simulated virtual ants. These virtual ants differ than the real world ants in two aspects.
	\begin{enumerate}
		\item Artificial ants have a prior knowledge of the distances between cities and therefore tend to select the shorter distance.
		\item Artificial ants have memory, they know which cities they have visited along a path and will not select those cities again.
	\end{enumerate}
	We can express the probability of ant \textit{k} selecting city \textit{j} after city \textit{i} as follows:

@@

	\section{Results}
	The codes to solve the TSP problem [links!!!] were executed on both Matlab and Octave, taking into consideration the resulting output path length and the execution time. All the test conducted were repeated ten times each in order to see how well the algorithms can improve in both criteria. In the program using GA, the running time on Octave was 45,362 seconds (approximately 12.5 hours). Due to the long duration of the test runtime this experiment was not repeated except two times. While the running time on Matlab provided a minimum time of 76 seconds on the first trial and a maximum time of 118.9 seconds on the 8$^th$ trial, the shortest path found was on the 6$^th$ iteration, of length 226,854.7 which is 65\% within the optimal solution. While the optimized GA algorithms produced much better running times for the same number of iterations with a minimum of 23.8 seconds on th first run and a maximum of 32.1 seconds on the 7$^th$ run, the final path length didn't differ much from the previous implementation, a path of length 200068 was found which is 60\% within the optimal solution. At a much faster running time and a slight improvement of 5\%, the seconds GA algorithm can be run for longer iterations and thus providing much better results.

@@

	The nearest neighbor tests conducted on Octave provided running times ranging between 132 seconds and 136 seconds, which are approximately 7 to 10 times less than all other conducted tests of different algorithms. On Matlab, the tests took 10.8 seconds for the quickest run and 15.4 seconds for the longest run on the 1$^st$ and 5$^th$ runs respectively. The nearest neighbor implementation conducted did not take a random starting point and therefore all runs provided the same path length of 96,514.8 which is approximately 18\%  within the optimal solution.

@@

@@

>>>>

\usepackage{bm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{url}
\usepackage{multirow}

\begin{document}	
	\twocolumn[
	\begin{@twocolumnfalse}
		\author{Mohamed Moanis Ali, Michael Sedrak \\  mohamed.moanis.ali@gmail.com, m.sedrak@live.com}
		\title{Solving TSP With Genetic Algorithms}
		\maketitle
		\begin{abstract}
			Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse condimentum, urna nec luctus tincidunt, nisi augue vehicula nulla, sit amet sollicitudin nisi nibh eget diam. Vivamus maximus tortor dolor, non porttitor urna vestibulum eu. Nullam maximus in massa eu lacinia. Curabitur ac risus ac elit euismod placerat quis eu odio. Sed ante ex, congue sit amet enim et, aliquam egestas turpis. Cras purus ex, viverra sit amet augue id, tempus ultricies risus. Aenean eget vestibulum orci. Ut enim nisi, tempor id urna et, dignissim ornare lectus. Nunc lobortis risus neque, a fermentum ante consequat ac. Fusce mollis pretium pellentesque. Vestibulum fermentum ornare elementum. Nam rutrum sem ut est mollis consequat. Morbi at faucibus mi. In ultricies justo non lorem varius, non tristique ante suscipit. Curabitur eget diam tincidunt, venenatis enim sagittis, malesuada magna. Nullam in urna a odio finibus tristique.
		\end{abstract}
	\end{@twocolumnfalse}
	]

@@

	\subsection{Nearest Neighbor (NN)}
	Nearest neighbor approach is one of the simplest algorithms in combinatorial optimization problems and also one of the fastest running algorithms\cite{GreedyNN}.

@@

	\begin{figure}[h!]
		\centering
		\includegraphics[scale=0.8]{./ACOexample.jpg}
		\caption{ACO example\cite{ACOExample} (a)The initial description of the problem showing distances. (b)At time t=0 no trail is found on the graph therefore ants choose the paths with equal probability. (c)Trails from previous step are found and therefore the choice of paths is biased towards the shorter edges.}
		\label{fig:ACOexample}
	\end{figure}

	For better understanding of the ACO algorithms, an example from\cite{ACOExample} is illustrated. As shown in Fig.\ref{fig:ACOexample} a, the system is shown with the initial start points A as the food source ,and E as the colony nest, and the distances of the paths. At t=0, 30 ants start moving from both E and A simultaneously. The ants currently can see no preference between the different paths because there are no traces of pheromone left behind, therefore, the ants divide randomly with equal probability going in to both points H and C, i.e. 15 ants are going from D to C and 15 ants from D to H, and similarly from point B to both C and H. At t=1, the 30 new ants coming from B, find 2 unequal paths, the path going to H has a trail of 15 from the ants that went to H from B, while the path going to C has a trail of 30 that can be divided into:
	\begin{enumerate}
		\item The trail from ants going from point B.
		\item The trail from ants coming from point D.
	\end{enumerate}
	 Resulting in a bias in the probability of choosing the path unlike the previous step. The number of ants going to C is expected to be double the number of ants going to H. The same is true for the 30 new ants coming from D.

	In combinatorial problems\cite{ACO}, we deal with simulated virtual ants. These virtual ants differ than the real world ants in two aspects.
	\begin{enumerate}
		\item Artificial ants have a prior knowledge of the distances between cities and therefore tend to select the shorter distance.
		\item Artificial ants have memory, they know which cities they have visited along a path and will not select those cities again.
	\end{enumerate}
	We can express the probability of ant \textit{k} selecting city \textit{j} after city \textit{i} as follows:

@@

	\section{Results}
	The codes to solve the TSP problem [links!!!] were executed on both Matlab and Octave, taking into consideration the resulting output path length and the execution time. All the test conducted were repeated ten times each in order to see how well the algorithms can improve in both criteria. In the program using GA, the running time on Octave was 45,362 seconds (approximately 12.5 hours). Due to the long duration of the test runtime this experiment was not repeated except two times. While the running time on Matlab provided a minimum time of 76 seconds on the first trial and a maximum time of 118.9 seconds on the 8$^th$ trial, the shortest path found was on the 6$^th$ iteration, of length 226,854.7 which is 65\% within the optimal solution. While the optimized GA algorithms produced much better running times for the same number of iterations with a minimum of 23.8 seconds on th first run and a maximum of 32.1 seconds on the 7$^th$ run, the final path length didn't differ much from the previous implementation, a path of length 200068 was found which is 60\% within the optimal solution. At a much faster running time and a slight improvement of 5\%, the seconds GA algorithm can be run for longer iterations and thus providing much better results. Results of GA run is shown in fig.\ref{fig:GAalgorithm}

	\begin{figure}[h!]
		\centering
		\includegraphics[scale=0.25]{./GA1.jpg}
		\includegraphics[scale=0.25]{./GA2.jpg}
		\caption{TSP using GA algorithm}
		\label{fig:GAalgorithm}
	\end{figure}

@@

	The nearest neighbor tests conducted on Octave provided running times ranging between 132 seconds and 136 seconds, which are approximately 7 to 10 times less than all other conducted tests of different algorithms. On Matlab, the tests took 10.8 seconds for the quickest run and 15.4 seconds for the longest run on the 1$^st$ and 5$^th$ runs respectively. The nearest neighbor implementation conducted did not take a random starting point and therefore all runs provided the same path length of 96,514.8 which is approximately 18\%  within the optimal solution. Results of NN run is shown in fig.\ref{fig:NNalgorithm}

	\begin{figure}[h!]
		\centering
		\includegraphics[scale=0.25]{./NN.jpg}
		\caption{TSP using NN algorithm}
		\label{fig:NNalgorithm}
	\end{figure}

@@

	The previous results are tabulated in Table \ref{tab:results}

	\begin{table}[h!]
		\caption{Results of TSP algorithms}
		\scalebox{0.8}{
		\begin{tabular}{cccccc} 
			\hline
			{Algorithm} &{GA} &{SA} & {NN} & {ACO} &{OGA}\\

			\hline
			{\multirow{2}{*}{Octave}} & {45362} &{1217}& {132} &{777} &{N/A} \\
			 running time & & & & & \\

			\hline
			{\multirow{2}{*}{Matlab}} & {76} &{63}& {11} &{12} &{24} \\
			running time & & & & & \\

			\hline
			{\multirow{2}{*}{Generated}} & {226854} &{89237}& {96514} &{951387} &{200068} \\
			path length & & & & & \\

			\hline
			{\multirow{2}{*}{Efficiency}} & {65\%} &{11.3\%}& {18\%} &{91.5\%} &{60\%} \\
			of path & & & & & \\

			\hline
		\end{tabular}}
	\label{tab:results}
	\end{table}

@@

		\bibitem{GreedyNN}
		Bendall, G. and Margot, F., \emph{Greedy Type Resistance of Combinatorial Problems}. Discrete Optimization 3 (2006).
		\bibitem{ACOExample}
		M. Dorigo, V. Maniezzo, A. Colorni, \emph{Ant system: optimization by a colony of cooperating agents}.(1996).
		\bibitem{ACO}
		Jinhui Y., Xiaohu S., Maurizio M., Yanchun L.,\emph{An ant colony optimization method for generalized TSP problem}. (2008).
